---
title: "p8105_hw3_nr2774"
author: "Nergui"
date: "2022-10-09"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 1

```{r}
#devtools::install_github("p8105/p8105.datasets")
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(patchwork)
library(p8105.datasets)
library(lubridate)
data("instacart")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

### Problem 2 

```{r}
# Load, tidy, and otherwise wrangle the data.

accel_df = read_csv("data_hw3/accel_data.csv",
                    col_types = cols(
      week = col_integer(),
      day_id = col_integer()
      )) |> 
  janitor::clean_names() |> 
    mutate(
    weekend = case_when(
      day == "Monday" ~ "weekday",
      day == "Tuesday" ~ "weekday",
      day == "Wednesday" ~ "weekday",
      day == "Thursday" ~ "weekday",
      day == "Friday" ~ "weekday",
      day == "Saturday" ~ "weekend",
      day == "Sunday" ~ "weekend"
    ),
    day = fct_relevel(day,"Monday", "Tuesday", "Wednesday", "Thursday",
                      "Friday", "Saturday", "Sunday"))

accel_df_pivot = accel_df |> 
  relocate("weekend") |> 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity",
    names_prefix = "activity_",
    values_to = "activity_count") |> 
  rename(activity_min = activity) |> 
  mutate(activity_min = as.numeric(activity_min))

```

* There are `r nrow(accel_df)` observations in the activiy dataset along with  `r ncol(accel_df)` variables. Those variables are called week, day_id, activity_1:activity_1440. 

```{r}
# create a table
accel_df_pivot_min = 
  accel_df_pivot |> 
  group_by(day_id, day, week) |> 
  summarize(total_activity = sum(activity_count),
                        .groups = "drop") |> 
    ungroup()  


accel_df_pivot_min |> 
  select(-day_id) |>
  pivot_wider(
    names_from = "day",
    values_from = "total_activity")

```


* Activity counts were lowest on the Saturdays of (weeks 4 and 5). The individual appears to be most active(gradualing increasing activity count) on weeks days. 

```{r}
## Single - panel plot for 24-hours activity

accel_df_pivot |> 
  ggplot(aes(x = activity_min, y = activity_count, color = day)) + geom_point() + 
  labs(
    title = "24-Hour Activity Count by Day",
    x = "Time",
    y = "Activity Count",
    caption = "Data from the accel dataset")
```


### Problem3

```{r}
#Exploration of data 

library(p8105.datasets)
data("ny_noaa")

```

The NY_NOAA dataset was driven from NOAA National Climatic Data Center. It contains information from all New York state weather stations from January 1, 1981 through December 31, 2010.
The NOAA dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns.

The variables in the data set are:

* `id`: Weather station ID
* `date`: Date of observation
* `prcp`: Precipitation (tenths of mm)
* `snow`: Snowfall (mm)
* `snwd`: Snow depth (mm)
* `tmax`: Maximum temperature (tenths of degrees C)
* `tmin`: Minimum temperature (tenths of degrees C)

There are `r count(distinct(ny_noaa, id))` distinct weather stations in NYS.

Because not all weather stations in NY collect all these variables, the original dataset contains extensive amount of missing data.

To take a closer look at the amount of missing data, let's look at what proportion of data is missing for each variable.

```{r}
noaa_missing = ny_noaa |> 
  summarize(
    missing_prcp = mean(is.na(prcp)),
    missing_snow = mean(is.na(snow)),
    missing_snwd = mean(is.na(snwd)),
    missing_tmax = mean(is.na(tmax)),
    missing_tmin = mean(is.na(tmin))
  ) |> 
  knitr::kable()
```

* Data shows that there nearly 50% of data is missing in `tmax` and `tmin`. 43 % for each of these variables have missing data. 

```{r}
# Data cleaning ~ Tidying 
noaa_tidy = ny_noaa |>  
  janitor::clean_names() |>  
  separate(date, into = c("year", "month", "day"), convert = TRUE) |> 
  mutate(
    prcp = prcp / 10 ,
    tmax = as.numeric(tmax) / 10,
    tmin = as.numeric(tmin) / 10,
    year = as.numeric(year),
    day = as.numeric(day), 
    month = recode_factor(month,
          "01" = "January",
          "02" = "February",
          "03" = "March",
          "04" = "April",
          "05" = "May",
          "06" = "June",
          "07" = "July",
          "08" = "August",
          "09" = "September",
          "10" = "October",
          "11" = "November",
          "12" = "December"
          )) |>  
  relocate(year, month, day, everything())
```
  

```{r}
noaa_tidy |>  
  count(snow) |> 
  arrange(desc(n))
```

* For snowfall, the most commonly observed value is '0', indicating that the most days in NY is not snowy. 


```{r}
# Two-panel plot ~ Average Max Temperature 
tmax_noaa = noaa_tidy |>  
  filter(month %in% c("January", "July")) %>% 
  group_by(id, year, month) |>  
  summarize(mean_tmax = mean(tmax, na.rm = TRUE))
ggplot(tmax_noaa, aes(x = year, y = mean_tmax, color = month)) +
  geom_point(alpha = 0.2) +
  geom_smooth(alpha = 0.5) +
  labs(
    title = "Avg Max Temp in January and July in each NY Station Across Years",
    x = "Year",
    y = "Temp (°C)",
    caption = "Data from the NOAA dataset"
  ) +
  scale_x_continuous(
    breaks = c(1980, 1985, 1990, 1995, 2000, 2005, 2010)
    ) +
  facet_grid(. ~ month) +
  scale_color_manual(values = c("navy", "orange")) +
  theme(
    legend.position = "none")

```

*The average max temp in January is usually within the -10-10°C range. There was not much difference in average max temperatures between 1980 and 2010. There are 2 sinfiifant outlier observed in years of 1980 and 2005. 

geom_sooth indicated that there might a slight increase in January over years when its compared to 1980 and 2010. By running statistical analysis, we will know the significance level. 

Average max temperatures in July were mostly bewteen 25-30°C range. There were couple outliers. However, tempearture difference over the years were no apparently different. 

```{r}
# Two panel plot ~ tmax vs tmin 
tmax_tmin_plot = ggplot(noaa_tidy, aes(x = tmin, y = tmax)) +
  geom_hex(bins = 40) +
  geom_smooth(se = F)
  labs(
    title = "NY Temperatures (1981-2010)",
    x = "Min Temp (°C)",
    y = "Max Temp (°C)"
  ) + 
  theme(legend.position = "right")
tmax_tmin_plot

```


```{r}
library(ggridges)
snowfall_plot = noaa_tidy |>  
  filter(snow > 0, snow < 100) |>  
  ggplot(aes(x = snow, y = factor(year), fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) + 
  scale_fill_viridis_c(name = "Snowfall (mm)", option = "A") + 
  labs(
    title = "NY Snowfall Values, 0-100mm (1981-2010)",
    x = "Snowfall (mm)",
    y = "Year"
  )
snowfall_plot
```

